---
title: "Prosodic Fingerprint"
output:
  html_document:
    df_print: paged
---

```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
library(tidyverse)
library(plyr)
library(ggplot2)
library(effects)
library(lmerTest)
library(Hmisc)
```


## Data preprocessing
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Unigram_data <- read_csv2("Unigrams.csv")
Unigram_data <- Unigram_data[-which(Unigram_data$Speaker_ID == "Speaker_ID"),]
nrow(Unigram_data)
Unigram_data$Time_s <- as.double(Unigram_data$Time_s)
Unigram_data$Abs_value <- as.double(Unigram_data$Abs_value)
Unigram_data$Age <- as.double(Unigram_data$Age)
Unigram_data$Sex <- as.factor(Unigram_data$Sex)
#The functions define the boundaries of unigrams
cut_unigram <- function(Abs_value){
  unigrams <- cut(Abs_value, breaks = 3, labels = c(-1,0,1))
  return(unigrams)
}
cut_unigram_raw <- function(Abs_value){
  unigrams <- cut(Abs_value, breaks = 3)
  return(unigrams)
}

# Ordering the data by speaker
Unigram_data <- Unigram_data[order(Unigram_data$Speaker_ID),]
# Normalizing (Z-scaling the) absolute pitch value 
Unigram_data$zAbs_value <- ave(as.numeric(Unigram_data$Abs_value), Unigram_data$Speaker_ID, FUN=scale)
# Cutting the unigrams into intervals
zUnigram_cut <- with(Unigram_data, tapply(zAbs_value, Speaker_ID,cut_unigram_raw))
zUnigram_raw <- c()

#Attaching normalized dataand intervals to the other data
for (element in zUnigram_cut){
  zUnigram_raw <- c(zUnigram_raw,as.character(element))
}
Unigram_data <- cbind(Unigram_data,zUnigram_raw)

#Counting z-normalized unigrams
zUnigram_cut <- with(Unigram_data, tapply(zAbs_value, Speaker_ID,cut_unigram))
zUnigram <- c()
for (element in zUnigram_cut){
  zUnigram <- c(zUnigram,as.character(element))
}
Unigram_data <- cbind(Unigram_data,zUnigram)
Unigram_data$Unigram <- as.factor(Unigram_data$Unigram)
Unigram_data$zUnigram <- as.factor(Unigram_data$zUnigram)
#levels(Unigram_data$zUnigram)

#Calculating deltas and normalized deltas
zDelta <- diff(as.numeric(zUnigram))
zDelta <- c(NA,zDelta)
Unigram_data <- cbind(Unigram_data,zDelta)
Delta <- diff(as.numeric(Unigram_data$Unigram))
Delta <- c(NA,Delta)

#attaching the deltas to the data
Unigram_data <- cbind(Unigram_data,Delta)
Unigram_data$Delta <- as.factor(Unigram_data$Delta)
Unigram_data$zDelta <- as.factor(Unigram_data$zDelta)

# Creating unique ID's for each sentence
Unigram_data <- cbind(paste(Unigram_data$Speaker_ID,Unigram_data$Sentence),Unigram_data)
colnames(Unigram_data)[1] <- "ID"
Unigram_data$ID <- as.factor(Unigram_data$ID)
#Unigram_data$Unigram <- as.factor(Unigram_data$Unigram)

# Counting raw unigrams
df <- with(Unigram_data, tapply(Unigram, ID, plyr::count))
Raw_Ug_Counted <- ldply(df, data.frame)
Raw_Ug_Counted <- spread(Raw_Ug_Counted, x, freq)
colnames(Raw_Ug_Counted)[1] <- "ID"

Raw_Ug_Counted$Speaker_ID <- Unigram_data$Speaker_ID[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Ug_Counted$Sex <- Unigram_data$Sex[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Ug_Counted$Age <- Unigram_data$Age[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Ug_Counted$Place <- Unigram_data$Place[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Ug_Counted$Text <- Unigram_data$Text[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Ug_Counted$Sentence <- Unigram_data$Sentence[match(Raw_Ug_Counted$ID, Unigram_data$ID)]

#Counting raw deltas
bdf <- with(Unigram_data, tapply(Delta, ID, plyr::count), default = 0)
Raw_Dlt_Counted <- ldply(bdf, data.frame)
Raw_Dlt_Counted <- spread(Raw_Dlt_Counted, x, freq)
colnames(Raw_Dlt_Counted)[1] <- "ID"

Raw_Dlt_Counted$Speaker_ID <- Unigram_data$Speaker_ID[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Dlt_Counted$Sex <- Unigram_data$Sex[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Dlt_Counted$Age <- Unigram_data$Age[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Dlt_Counted$Place <- Unigram_data$Place[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Dlt_Counted$Text <- Unigram_data$Text[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
Raw_Dlt_Counted$Sentence <- Unigram_data$Sentence[match(Raw_Ug_Counted$ID, Unigram_data$ID)]

#Counting z-normalised Unigrams
zdf <- with(Unigram_data, tapply(zUnigram, ID, plyr::count))
zUg_Counted <- ldply(zdf, data.frame)
zUg_Counted <- spread(zUg_Counted, x, freq)
zUg_Counted[is.na(zUg_Counted)] <- 0
colnames(zUg_Counted)[1] <- "ID"

zUg_Counted$Speaker_ID <- Unigram_data$Speaker_ID[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zUg_Counted$Sex <- Unigram_data$Sex[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zUg_Counted$Age <- Unigram_data$Age[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zUg_Counted$Place <- Unigram_data$Place[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zUg_Counted$Text <- Unigram_data$Text[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zUg_Counted$Sentence <- Unigram_data$Sentence[match(Raw_Ug_Counted$ID, Unigram_data$ID)]

#Counting z-normalized Deltas
zbdf <- with(Unigram_data, tapply(zDelta, ID, plyr::count))
zDlt_Counted <- ldply(zbdf, data.frame)
zDlt_Counted <- spread(zDlt_Counted, x, freq)
zDlt_Counted[is.na(zDlt_Counted)] <- 0
colnames(zDlt_Counted)[1] <- "ID"

zDlt_Counted$Speaker_ID <- Unigram_data$Speaker_ID[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zDlt_Counted$Sex <- Unigram_data$Sex[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zDlt_Counted$Age <- Unigram_data$Age[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zDlt_Counted$Place <- Unigram_data$Place[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zDlt_Counted$Text <- Unigram_data$Text[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
zDlt_Counted$Sentence <- Unigram_data$Sentence[match(Raw_Ug_Counted$ID, Unigram_data$ID)]
```

## Exploratory plots

Figure 1. Pitch sub-ranges in a recording sample
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Unigram_data$Time_s <- as.numeric(Unigram_data$Time_s)
Unigram_data$Abs_value <- as.numeric(Unigram_data$Abs_value)
Unigram_data$ID <- as.character(Unigram_data$ID)
subset(Unigram_data, Unigram_data$ID %in% c("AnnaSh 2")) %>% 
  ggplot(aes(Time_s, Abs_value))+
  xlab("Time")+
  ylab("Pitch Value")+
  geom_hline(aes(yintercept=85.56109814999999), color = "red")+
  geom_hline(aes(yintercept=220.06107974516), color = "blue")+
  geom_hline(aes(yintercept=354.56106134032), color = "blue")+
  geom_hline(aes(yintercept=489.06104293548003), color = "red")+
  annotate(geom="text", x = 1, 
           y = 75.56109814999999, label = "Outliers", colour='red', size = 3) +
  annotate(geom="text", x = 1, 
           y = 150, label = "-1", colour='blue', size = 4) +
  annotate(geom="text", x = 1, 
           y = 300, label = "0", colour='blue', size = 4) +
  annotate(geom="text", x = 1, 
           y = 450, label = "1", colour='blue', size = 4) +
  annotate(geom="text", x = 1, 
           y = 498.06104293548003, label = "Outliers", colour='red', size = 3) +
  #xlim(min(Unigram_data[Unigram_data$ID == "AnnaSh 10",]$Time_s), max(Unigram_data[Unigram_data$ID == "AnnaSh 10",]$Time_s))+
  xlim(93.5,96.5)+
  geom_line()+
  #facet_wrap(. ~ ID)+
  theme_bw() -> Fig1
```

```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Fig1
```

### Figure 2 (corrected age groups). The Distribution of the z-scored Unigram Values by Place, Sex and Age

After the text was published, I noticed a small mistake in the group names. The age groups should be named "≤40" instead of "25-40" and ">40" instead of "45+". The mistake follows from the data collection protocol that we developed at some point. Since the current case study is partly based on the data collected before the protocol was established, the speakers could not be divided into such groups and I used 40 years as the cut-off point, but the plot titles and the description in the text were not changed accordingly. The mistake does not affect the counts and modelling. This grouping is only used for convenience in barplots, where age cannot be treated as a numeric variable; the models treat in as numeric. In the two plots below, the mistake has been corrected.

```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Unigram_data <- mutate(Unigram_data, Age_Group = as.factor(ifelse(Unigram_data$Age <= 40,"25-40 (corrected: ≤ 40 )","45+ (corrected: >40)")))
Unigram_data$Place_1 <- factor(Unigram_data$Place, levels = c("Krasnoyarsk", "Nakhodka", "Novosibirsk", "Moscow" ))
subset(Unigram_data, !is.na(zUnigram)) %>% 
  ggplot(aes(zUnigram, color=Place, fill=Place))+
  geom_bar(stat = "count", position="dodge")+
  xlab("Distribution of unigrams by Place, Sex and Age")+
  ylab("Number of unigrams of each type")+
  facet_wrap(~Place_1:Sex:Age_Group) -> Fig2
```

```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Fig2
```

### Figure 3 (corrected age groups). The Distribution of the z-scored Delta Values by Place, Sex and Age
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
zhist <- subset(Unigram_data, !is.na(zDelta))
zhist <- zhist[zhist$zDelta != 0,]
subset(zhist, !is.na(zDelta)) %>% 
  ggplot(aes(zDelta, color=Place, fill=Place))+
  geom_bar(stat = "count", position="dodge")+
  xlab("Distribution of deltas by Place, Sex and Age")+
  ylab("Number of deltas of each type")+
  facet_wrap(~Place_1:Sex:Age_Group) -> Fig3
```

```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Fig3
```




### Adjusting the unigram data for modelling
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
data_ug <- zUg_Counted
row.names(data_ug) <- paste(data_ug$Speaker_ID,data_ug$Sentence)
data_ug$X1prop <- data_ug$"-1"/(data_ug$"0" + data_ug$"1" + data_ug$"-1" + 1)
data_ug <- mutate(data_ug, TextType = as.factor(ifelse(grepl("Experiment",Text),"dialogue","monologue")))
data_ug <- mutate(data_ug, Role = as.factor(ifelse(grepl("Follower",Text),"hearer","speaker")))
data_ug$Age_Group<-c("Low", "High")[
  findInterval(data_ug$Age , c(-Inf, 40, Inf) ) ]
data_ug$Text <- as.factor(data_ug$Text)
data_ug$Sex <- as.factor(data_ug$Sex)
data_ug$Age_Group <- as.factor(data_ug$Age_Group)
```

### The same for deltas
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
data_dlt <- zDlt_Counted
row.names(data_dlt) <- paste(data_dlt$Speaker_ID,data_dlt$Sentence)
data_dlt$X0prop <- (data_dlt$"0")/(data_dlt$"1" + data_dlt$"-1" + data_dlt$"2" + data_dlt$"-2" + data_dlt$"0" + 1)
data_dlt <- mutate(data_dlt, TextType = as.factor(ifelse(grepl("Experiment",Text),"dialogue","monologue")))
data_dlt <- mutate(data_dlt, Role = as.factor(ifelse(grepl("Follower",Text),"hearer","speaker")))
data_dlt$Age_Group<-c("Low", "High")[
  findInterval(data_dlt$Age , c(-Inf, 40, Inf) ) ]
data_dlt$Text <- as.factor(data_dlt$Text)
data_dlt$Sex <- as.factor(data_dlt$Sex)
data_dlt$Age_Group <- as.factor(data_dlt$Age_Group)
```

## Modelling

### Model for -1 vs. other unigrams

```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Data_counted <- data_ug
summary(model.0 <- lmer(X1prop ~ Sex*Age + Place + TextType + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.0, test = "Chisq") # The Sex:Age interaction can be dropped
summary(model.1 <- lmer(X1prop ~ Sex + Age + Place + TextType + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.1, test = "Chisq") # Place can be dropped
summary(model.2 <- lmer(X1prop ~ Sex + Age + TextType + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.2, test = "Chisq") # Age can be dropped
anova(model.1, model.2) # No significant difference between the models
summary(model.3 <- lmer(X1prop ~ Sex + TextType + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.3, test = "Chisq") # Nothing can be dropped, still test against a simpler model
summary(model.4 <- lmer(X1prop ~ Sex + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.4, test = "Chisq")
anova(model.3, model.4) # Model 3 is significantly better, so use it as the final model
```

### Figure 4. The effect of biological sex and text type on the proportion of “−1” unigrams to “not −1” unigrams
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
plot(allEffects(model.3))
```

### Model for 0 vs. other deltas
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Data_counted <- data_dlt
summary(model.0 <- lmer(X0prop ~ Sex*Age + Place + Role + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.0) # Place can be dropped
summary(model.1 <- lmer(X0prop ~ Sex*Age + Role + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.1) # Role can be dropped
summary(model.2 <- lmer(X0prop ~ Sex*Age + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.2) # The interaction is on the edge of significance, but still drop
anova(model.1, model.2) # No significant difference between the models, select the simpler one
summary(model.3 <- lmer(X0prop ~ Sex + Age + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.3) # Age can be dropped
summary(model.4 <- lmer(X0prop ~ Sex + (1 | Speaker_ID), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.4) # Nothing can be dropped
```

### Figure 5. The effects of biological sex and age on the proportion of “0” and “not 0” deltas
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
plot(allEffects(model.4))
```

### Model for -1 vs. other unigrams by sex and age and type of text, with place as random effect
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Data_counted <- data_ug
summary(model.0 <- lmer(X1prop ~ Sex*Age + TextType + (1 | Place), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.0, test = "Chisq") # Can't drop anything, all effects significant
```

### Figure 6. Effect of age by sex and text type on the proportion of “−1” unigrams
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
plot(allEffects(model.0))
```

### Model for 0 vs. other deltas by sex and age and type of text, with place as random effect

```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Data_counted <- data_dlt
summary(model.0 <- lmer(X0prop ~ Sex*Age + TextType + (1 | Place), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.0, test = "Chisq")
summary(model.1 <- lmer(X0prop ~ Sex*Age + (1 | Place), data = Data_counted, control = lmerControl(optimizer = "bobyqa")))
drop1(model.1, test = "Chisq")
summary(model.2 <- lm(X0prop ~ Sex*Age, data = Data_counted))
AIC(model.1)
AIC(model.2) #AIC of Model 2 is greater, select Model 1
```

### Figure7. Effect of age by sex on the proportion of “0” deltas
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
plot(allEffects(model.1))
```

### Interaction Models

In the next two moels we do not run model selection, because we're interested in the top-level interaction. 
Instead, we simply explore the significance levels of the effects and their combinations.

```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Data_counted <- data_dlt
Data_counted <- Data_counted[Data_counted$Place %nin% c("Moscow"),]
Data_counted$Speaker_ID <- factor(Data_counted$Speaker_ID)
Data_counted$Place <- factor(Data_counted$Place)
summary(model.0 <- lm(X0prop ~ Place:Age:Sex, data = Data_counted))
```

### Figure 8. Effect of age and sex by region on the proportion of “−1” unigrams
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
plot(allEffects(model.0))
```

```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
Data_counted <- data_ug
Data_counted <- Data_counted[Data_counted$Place %nin% c("Moscow"),]
Data_counted$Speaker_ID <- factor(Data_counted$Speaker_ID)
Data_counted$Place <- factor(Data_counted$Place)
summary(model.0 <- lm(X1prop ~ Place:Sex:Age, data = Data_counted))
```

### Figure 9. Effect of age and sex by region on the proportion of “0” deltas
```{r message = FALSE, error = FALSE, warning = FALSE, fig.width=20,fig.height=10}
plot(allEffects(model.0))
```

